---
title: "In-class Exercise 3: Spatial Point Patterns Analysis"
author: "Pan Mingwei"
date: "September 2, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Installing maptools

maptools is retired and binary is remove from CRAN. However, we can download from Posit Public Package Manager snapshot using the code chunk below:

```{r}
#| eval: false
install.packages("maptools",
                 repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

::: callout-note
To avoid install `maptools` every time the page being render, add "**#\| eval: false**" in the code chunk.
:::

# Creating coastal outline

## Working with st_union()

```{r}
pacman::p_load(sf,tidyverse)
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

```{r}
plot(sg_sf)
```

::: callout-note
`owin` is a window specifies the region of space within which spatial points are observed or analyzed. Represented a domain or boundary of the spatial data.
:::

# Issue encounter in hands-on exercise

The problem of this is because `maptools` is not installed.

```{r}
#| eval: false
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

# Import data

First convert the data frame into sf object, and transform the projection for Myanmar and convert the `event_date` field data type from Char to Date.

```{r}
acled_sf <- read_csv("data/ACLED_Myanmar.csv") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(event_date = dmy(event_date))
```
