---
title: "Hands-on Exercise 1"
author: "Pan Mingwei"
---

## **1.3 Getting Started**

Install and launching R packages

The code chunk below use p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(sf, tidyverse)
```

## 1.4 Importing the data 

#### 1.4.1 Importing polygon feature data in shapefile format

```{r}
mpsz = st_read(dsn = "/Users/mingwei/Desktop/SMU/Y3S1/IS415/xXxPMWxXx/IS415-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial/MasterPlan2014SubzoneBoundaryWebSHP", layer = "MP14_SUBZONE_WEB_PL")
```

#### 1.4.2 Importing polyline feature data in shapefile form

```{r}
cyclingpath = st_read(dsn = "/Users/mingwei/Desktop/SMU/Y3S1/IS415/xXxPMWxXx/IS415-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial/CyclingPath_Jul2024", layer = "CyclingPathGazette")
```

#### 1.4.3 Importing GIS data in kml format

```{r}
preschool = st_read("/Users/mingwei/Desktop/SMU/Y3S1/IS415/xXxPMWxXx/IS415-GAA/Hands-on_Ex/Hands-on_Ex01/data/geospatial/PreSchoolsLocation.kml")
```

## 1.5 Checking the Content of A Simple Feature Data Frame

#### 1.5.1 Working with st_geometry()

```{r}
st_geometry(mpsz)
```

#### 1.5.2 Working with glimpse()

```{r}
glimpse(mpsz)
```

#### 1.5.3 Working with head()

```{r}
head(mpsz, n=5)
```

## 1.6 Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## 1.7 Working with Projection 

#### 1.7.1 Assigning EPSG code to simple feature data frame

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz,3414)
```

```{r}
st_crs(mpsz3414)
```

#### 1.7.2 Transforming the projection of preschool from wgs84 to svy21.

```{r}
preschool3414 <- st_transform(preschool,crs = 3414)
head(preschool3414, n=1)
```

## 1.8 Importing and Converting An Aspatial Data 

#### 1.8.1 Importing the aspatial data

```{r}
listings <- read_csv("/Users/mingwei/Desktop/SMU/Y3S1/IS415/xXxPMWxXx/IS415-GAA/Hands-on_Ex/Hands-on_Ex01/data//aspatial/listings.csv")
```

```{R}
list(listings) 
```

#### 1.8.2 Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)

glimpse(listings_sf)
```

## 1.9 Geoprocessing with sf package 1.9.1 Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

#### 1.9.2 Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## 1.10 Exploratory Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
